---
title: "Canopy Openess Index in R"
author: "Edson Silva-Júnior"
format: 
  html:
    embed-resources: true
    toc: true
    toc-depth: 5
    toc-expand: 1
    toc-location: left
    toc-title: Tópicos
    number-sections: false
    df-print: paged
    smooth-scroll: true
execute: 
  cache: false 
editor: visual
theme: 
  light: flatly
  dark: darkly
---

```{=html}
<style>
body {
text-align: justify;
font-size: 20px}
blockquote {
  background-color: #FAEFA6;
  padding: 10px;
  font-size: 14.5px
}
</style>
```

# Canopy Openess Index in R

In ecology reasearchs about forest species, such leaf litter animals, plants and fungi, one useful environmental process is the canopy openness, which is the how much the tree canopy is open. This affects several ecological process, due the solar radiation rates achieving that ecosystems. For that, ecology researchers use fish eye lens to photograph the forest canopy, to better analyse those processes. Canopy may be measure by an index, where:

-   0: a canopy fully close;

-   1: there is not canopy.

Several image analysis softwares can generate canopy openness index analysis, such as ImageJ. Although, for many images, programming languages can automatize process, accelerating the research's progress. In R, it is possible to analyse a set of canopy images and getting its canopy openness indexes.

# Packages

For those analysis, we gonna use the packages:

-   terra: importing images as `spatRast` objets and get its propieties;

-   tidyverse: to assistant plots and loop processes;

-   tidyterra: to better plot the `spatRast` objets;

-   hemispheR: to binarize canopy imagens, for canopy openness indexes analysis.

Remind to confer whether the packages where previously installed to library them.

```{r}
library(terra)

library(tidyverse)

library(tidyterra)

library(hemispheR)
```

# Data

## Importing

Our files are images taken by fish eye lens. Those images are archived on `cropped-images` folder. Initialy, using `list.files()` function, we get imagens path link.

```{r}
images <- list.files(path = "cropped-images", 
                     pattern = ".png")

images
```

## Visualizing

The next step is the importing and visualizing the images.

To importing, we use `terra::rast()` function, loading images as `spatRast` objects. Then, we use `ggplot2` and `tidyterra` functions to plot. Moreover, we construct a loop for the every imagens, constructing a function and loading it with `purrr` package.

```{r, warning=FALSE}
visualizing_canopy <- function(x){
  
  x <- paste0("cropped-images/", x)
  
  raster_bi <- terra::rast(x)
  
  ggplots <- ggplot() +
    tidyterra::geom_spatraster_rgb(data = raster_bi) +
    scale_fill_continuous(na.value = "transparent") +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) 
  
  print(ggplots)
  
}

purrr::walk(images, visualizing_canopy)
```

# Calculating Canopy Openness

## Visualizing

To calculating canopy openness index, we first binarize images, to pixels equivalent to canopy tree equals to 0 value and pixels equivalent to open sky equals to 1 value. For that, we use `hemispheR::import_fisheye()` to crop images to circle shape and then `hemispheR::binarize_fisheye()` to binarize circle shaped images.

```{r}
canopy_visualizing <- function(x){ 
  
  x <- paste0("cropped-images/", x)
  
  analy <- stringr::str_glue("analysis for {x}") 
  
  file <- x  |>
    hemispheR::import_fisheye()  |>
    hemispheR::binarize_fisheye()

  ggplt <- ggplot() +
    tidyterra::geom_spatraster(data = file) +
    scale_fill_viridis_c(na.value = "transparent", breaks = seq(0, 1, 1)) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    labs(title = analy) +
    theme_bw()
  
  print(ggplt)
  
}

purrr::walk(images, canopy_visualizing)
```

## Calculating

Finally, with binarize raster, we measuring the ration pixels = 1 (open sky) count by all pixels image count. For that, we use the `terra::ncell()` function.

```{r, message = FALSE}
canopy_Openess <- function(x){ 
  
  x <- paste0("cropped-images/", x)
  
  stringr::str_glue("analysis for {x}") |> message()
  
  raster <- x |>
    hemispheR::import_fisheye() |>
    hemispheR::binarize_fisheye()
  
  values_1 <- raster[raster > 0] |> terra::ncell()
  
  values_all <- raster |> terra::ncell()
  
  result <- values_1 / values_all
  
  print(result)
  
}

purrr::walk(images, canopy_Openess)
```

Notice the images with more open sky presented higher canopy openness index values.
