---
title: "Canopy Openness Index Analysis with R"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Analysing Canopy Openness Index Analysis with R

Canopy openness index (COI) is often used as predictor variable in ecological reseaches. COI can be calculated for the ratio between the image pixels count understanded as the skt and the imagel total pixels count, ranging from 0 (a full closed canopy) to 1 (a full open canopy). Usely, it's used fisheye images.

$$
COI = \frac{N-Pixels_{sky}}{N-Pixels_{total}}
$$

initialy, canopy images are binarized, to posteriorly pixels be counted. In R, we can calculate COI easily by simples image analyses.

# Required Packages

For our analysis, we use the required packages:

- [terra](https://github.com/rspatial/terra): for import images as rasters to calculate raster pixels count;

- [tidyverse](https://www.tidyverse.org/packages): for transform data, generate visualizations by ggplot graphs and make loops, throught [purrr package](https://purrr.tidyverse.org);

- [tidyterra](https://dieghernan.github.io/tidyterra): for visualizing rasters;

- [hemispheR](https://canopyphotography.wordpress.com/2022/04/05/hemispher-an-r-package-for-fisheye-canopy-image-analysis): to recorte fisheye images and binarize them;

```{r}
library(terra)

library(tidyverse)

library(tidyterra)

library(hemispheR)
```

# Data

Our data are images, taken by a fisheye lens.

## Images path  

Firs, we informate images files path, as `list.files()` function, informing its directory path (`path`) and its file class (`pattern`).

```{r}
images <- list.files(path = "cropped-images", 
                     pattern = ".png")

images
```

## Visualizing cannopy images through a looping

First to analysis, we have our files path, we build a [loop](https://stackoverflow.com/questions/74794193/how-to-use-the-purrr-package-in-r-instead-of-for-loop-to-iterate-over-indices) to visualize our images, making a function and run it on `purrr::walk()`, informating our image paths and maked function. That step is only to visualizing our images.

```{r}
visualizing_canopy <- function(x){ 
  
  x <- paste0("cropped-images/", x)
  
  raster_bi <- terra::rast(x)
  
  ggplots <- ggplot() +
    tidyterra::geom_spatraster_rgb(data = raster_bi) +
    scale_fill_continuous(na.value = "transparent") +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) 
  
  print(ggplots)
  
}

purrr::walk(images, visualizing_canopy)
```

# Calculating Canopy Openess index

Our next step is to calculate Canopy Openess index (COI). Initialy, we import the images, throught `hemispheR::import_fisheye()` function, and binarize images, throught `hemispheR::binarize_fisheye()`. [hemispheR package](https://canopyphotography.wordpress.com/2022/04/05/hemispher-an-r-package-for-fisheye-canopy-image-analysis) binarize images for the images light incidence index. For now, lets see binarized images. To facilite our work, lets make a loop, as previously made.

## Visualizing each binarized images  through a looping

```{r, warning = FALSE}
canopy_visualizing <- function(x){ 
  
  path_file <- paste0("cropped-images/", x) 
  
  image_name <- stringr::str_remove(x, ".png")
  
  analy <- stringr::str_glue("Binnarized image for {image_name}") 
  
  file <- path_file  |>
    hemispheR::import_fisheye()  |>
    hemispheR::binarize_fisheye()
  
  ggplt <- ggplot() +
    tidyterra::geom_spatraster(data = file) +
    scale_fill_viridis_c(na.value = "transparent", breaks = seq(0, 1, 1)) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    labs(title = analy) +
    theme_bw()
  
  print(ggplt)
  
}

purrr::walk(images, canopy_visualizing)
```

## Calculating opennes Index for each images through a looping

Finaly, we calculate COI. To count sky pixels and image total pixels, we use `terra::ncell()` function. TO only sky pixel, we filter image raster, as `raster[raster > 0]`. The results are values, ranging from 0 to 1.

```{r}
canopy_Openess <- function(x){ 
  
  path_file <- paste0("cropped-images/", x)
  
  image_name <- stringr::str_remove(x, ".png")
  
  stringr::str_glue("Cannopy Opennes Index for {image_name}:") |> 
    crayon::green() |> 
    message()
  
  raster <- path_file |>
    hemispheR::import_fisheye() |>
    hemispheR::binarize_fisheye()
  
  values_0 <- raster[raster > 0] |> 
    terra::ncell()
  
  values_all <- raster |> 
    terra::ncell()
  
  result <- values_0 / values_all
  
  print(result)
  
} 

purrr::walk(images, canopy_Openess)
```

